---
title: "p8105_hw1_sri2121"
author: "Sameera"
date: "2026-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
library(tidyverse)
library(ggplot2)
```


# Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

Details of the dataset:

The dataset is describing the various species of penguins found in various islands, their bill length, bill depth, flipper length, body mass and sex.

The number of different species available in this dataset is `r length(unique(penguins$species))`. They are: `r unique(penguins$species)`.

The number of rows in the dataset is `r nrow(penguins)` and the number of columns is `r ncol(penguins)`. 

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE))` mm.

Scatterplot of flipper length vs bill length:

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, 
                     colour = species)) + 
	geom_point()
```

Saving the plot image

```{r}
ggsave("penguins_plot.png")
```

# Problem 2

Create a data frame comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”

```{r}
set.seed(3)
df = 
  tibble(
    random_sample = rnorm(10),
    vec_logical = random_sample > 0,
    vec_character = c("apple", "banana", "broccoli", "kiwi", "grapes", "raspberry", "potato", "onion", "garlic", "ginger"),
    vec_factor = factor(sample(c("low", "medium", "high"), size = 10, replace = TRUE))
  )

df
```

The mean of random_sample is `r mean(df$random_sample)`. 

Converting variables from one type to another

```{r}
vec_logical_numeric = as.numeric(df$vec_logical)
vec_character_numeric = as.numeric(df$vec_character)
vec_factor_numeric = as.numeric(df$vec_factor)
```



