---
title: "p8105_hw1_sri2121"
author: "Sameera Reem"
date: "2026-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
library(tidyverse)
library(ggplot2)
```


# Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

Details of the dataset:

The dataset is describing the various species of penguins found in various islands, their bill length, bill depth, flipper length, body mass and sex.

The number of different species available in this dataset is `r length(unique(penguins$species))`. They are: `r unique(penguins$species)`.

The number of rows in the dataset is `r nrow(penguins)` and the number of columns is `r ncol(penguins)`. 

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE))` mm.

Scatterplot of flipper length vs bill length:

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, 
                     colour = species)) + 
	geom_point()
```

Saving the plot image

```{r}
ggsave("penguins_plot.png")
```

# Problem 2

```{r}
set.seed(3)
df = 
  tibble(
    random_sample = rnorm(10),
    vec_logical = random_sample > 0,
    vec_character = c("apple", "banana", "broccoli", "kiwi", "grapes", "raspberry", "potato", "onion", "garlic", "ginger"),
    vec_factor = factor(
      sample(
        c("low", "medium", "high"), 
        size = 10, 
        replace = TRUE))
  )

df
```

The mean of random_sample is `r mean(pull(df, random_sample))`.\
The mean of vec_logical is `r mean(pull(df, vec_logical))`.\
The mean of vec_character is `r mean(pull(df, vec_character))`.\
The mean of vec_factor is `r mean(pull(df, vec_factor))`.\

Here, we get the mean for random_sample and vec_logical but for vec_character and vec_factor, the output is NA.\
This is because random_sample is a numerical variable and vec_logical codes TRUE as 1 and FALSE as 0, so it gives out a numerical value to calculate mean from.\
This is not the case for vec_character and vec_factor. Since there's no numerical value attributed to either variables, it calculates mean as NA.


Converting variables from one type to another:

```{r}
vec_logical_numeric = as.numeric(df$vec_logical)
vec_character_numeric = as.numeric(df$vec_character)
vec_factor_numeric = as.numeric(df$vec_factor)
```
When we convert the logical, character and factor variable to numeric, the logical variable gives an output of either 0 or 1. This is because the logical vector TRUE is coded as 1 and FALSE is coded as 0 as already mentioned. The factor variable gives an output of 1, 2 or 3. This is because the factor levels are stored as integers. The character vector gives an output of NA, since there's no numerical value R can attribute it to.




